---
title: "Exploration of Cereal Data"
author: Jen McGregor
output:
  html_document: default
  pdf_document: default
---
## Setup


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
```


## The Data

```{r read_data}
cereals <- read_csv("https://www.dropbox.com/s/ll2c9drmlom1ony/cereals.csv?dl=1")
```

Documentation for dataset: https://www.kaggle.com/crawford/80-cereals/version/2

```{r check_data, eval = FALSE}
#### Edit this code to examine your dataset
cereals %>%
summary()
```
The data set has 16 variables (before adding cal_per_cup later). 

## Data Cleaning and Transformation

```{r}
#### clean the data
cereals %>%
  drop_na()
```


```{r summarize}
#### Write code to show the mean and median and sd of sugar content per serving of all cereals
cereals %>%
  summarize(across(c(sugars), list(mean,median,sd)))

```

```{r summarize2}
#### Write code to show the total calories of all cereals
cereals %>%
  summarize(sum(calories))
```


```{r mutate}
#### Write code to create the variable "cal_per_cup" here
cereals <- cereals %>%
  mutate(cal_per_cup=calories/cups)

cereals %>%
  summarize(cal_per_cup)
```

```{r filter_select,error=TRUE}
#### Write code to include only Kellogg brand cereals, and only relevant columns
cereals %>%
  select(name,mfr,cal_per_cup)
  filter(mfr=='K')
```


```{r arrange}
#### Write code to sort the dataset by calories per cup
cereals %>%
  arrange(cal_per_cup)
```

```{r wrangle}

#### Combine all steps into one pipeline
Kelloggs <- cereals %>%
  mutate(cal_per_cup=calories/cups) %>%
  select(name,mfr,cal_per_cup) %>%
  filter(mfr=='K') %>%
  arrange(desc(cal_per_cup))
plot_layout(ncol=1)
```

```{r}
###long join

cereals_longer <- cereals %>%
  select(name, mfr, protein, sugars) %>%
  pivot_longer(protein:sugars,
               names_to = "Nutrient",
               values_to = "Amount")


## Visualizing the data
#### Make a plot

cereals_longer %>%
  ggplot(aes(x = mfr, y = Amount, fill = Nutrient)) +
  geom_boxplot()
```



## Conclusion

I learned that certain manufacturers have wide spreads of sugars within their cereals while other manufacturers do not as wide of spreads of sugars. For the most part, spread of protein seems to be less in cereals than spread of sugars. 
