---
title: "Week3HW"
author: "Jen McGregor"
date: "1/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
```

# 1 Obtain counts and proportions for the categories (not the subcategories).

```{r}
#base
table(kickstarter$category);
prop.table(table(kickstarter$category));  
#tidy
kickstarter%>%
  group_by(category)%>%
  summarize(n=n())

kickstarter%>%
  group_by(category)%>%
  summarize(n=n())%>%
  mutate(prop=count/sum(count))
```

# 2 Create a contingency table for the categories by funding status. Find sample proportions for the status conditional on the category.
```{r}
#base
table(kickstarter$category, kickstarter$status);
prop.table(table(kickstarter$category, kickstarter$status), margin=1);
#tidy
kickstarter%>%
  group_by(category,status)%>%
  summarize(n=n())
```

# 3 Obtain summary statistics for the following variables: funded percentage, updates, and duration.

```{r}
#base
apply(kickstarter[,c("funded_percentage", "updates","duration")],2,mean);
#tidy
kickstarter %>%
  summarize(mean_FP=mean(funded_percentage),mean_UP=mean(updates),mean_D=mean(duration))
kickstarter %>%
  summarize(median_FP=median(funded_percentage),median_UP=median(updates),median_D=median(duration))
kickstarter %>%
  summarize(sd_FP=sd(funded_percentage),sd_UP=sd(updates),sd_D=sd(duration))
kickstarter %>%
  summarize(length_FP=length(funded_percentage),length_UP=length(updates),length_D=length(duration))
```
# 4 obtain summary statistics for the funded percentage by category.
```{r}
#base
tapply(kickstarter$funded_percentage, kickstarter$category, mean);
tapply(kickstarter$funded_percentage, kickstarter$category, median);
tapply(kickstarter$funded_percentage, kickstarter$category, sd);
tapply(kickstarter$funded_percentage, kickstarter$category, length);
#tidy
kickstarter%>%
  group_by(category)%>%
  summarize(mean_FP=mean(funded_percentage))
```

# 5  obtain summary statistics for the funded percentage by category and status.
```{r}
#base
tapply(kickstarter$funded_percentage, list(kickstarter$category, kickstarter$status), mean);
#tidy
kickstarter%>%
  group_by(category,status)%>%
  summarize(mean_FP=mean(funded_percentage))
```
# 6 Find the correlation between the funded percentage and updates.
```{r}
#base
cor(kickstarter$funded_percentage,kickstarter$updates)
#tidy
kickstarter%>%
  summarise(correlation=cor(funded_percentage,updates))
```
# 7 Find the correlation between the funded percentage and updates by category and status. Only do this using tidy functionality.
```{r}
kickstarter%>%
  group_by(category,status)%>%
  summarise(correlation=cor(funded_percentage,updates))
kickstarter %>% group_by(category,status) %>% summarize(cor(funded_percentage,updates));
```

```{r}
?boxplot
rb <- boxplot(decrease ~ treatment, data = OrchardSprays, col = "bisque")
summary(rb)
title("Comparing boxplot()s and non-robust mean +/- SD")
```
